---
title: "Getting started with `casemix`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{casemix-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

This vignette will demonstrate how to use the `casemix` package to nonparametrically estimate the case-mix adjusted quality of many units. A prominent application of these methods is estimating the quality of health care providers (e.g., hospitals, health insurance plans, doctors) adjusted for the mix of patients that they see. These methods are also used in education research where the units might be schools or teachers, and adjustment might be made for the mix of students taught by school/teacher. For the purposes of demonstration, we will use the terminology of providers (for units) and patients (for individual observations) throughout the rest of this demonstration.

There are two types of quality estimates generated by `casemix`: population-weighted quality estimates and equity-weighted quality estimates. Population-weighting estimates the quality of providers standardized to the population of all patients. These answer the counterfactual question of what the provider's quality would be if they saw a mix of patients similar to the population distribution of patients. Equity-weighting estimates quality standardized to a different population defined by certain equity features. Unlike population-weighted quality, which can be determined mostly by the largest groups of patients, equity-weighting ensures that certain groups of patients (defined by, e.g., race, sex, income, etc.) are equally represented in the quality measurement. 

## Preliminaries

First, we load a sample dataset and a dataset that indicates the true underlying quality of each provider. In the dataset, `id` is a random patient ID, `unit` indicates the provider, `y` indicates the binary measure of interest, `w1`, `w2`, `w3` are characteristics on which we might wish to ensure equity (e.g., race, sex, income) and `x1`, `x2`, `x3` are other characteristics that we wish to control for.

```{r setup}
library(casemix)
library(dplyr)

data(sample_dataset)
data(true_quality)
```

We can first look at the true quality for each of the ten providers

```{r}
knitr::kable(true_quality %>%
               arrange(desc(true_quality)), digits = 3)
```

We can also see the results we would obtain if we did not adjust for patient mix:

```{r}
knitr::kable(sample_dataset %>%
  group_by(unit) %>%
  summarise(unadjusted_quality = mean(y)), digits = 3)
```

## Population-weighted quality

If there are no particular equity concerns, but one only wishes to estimate provider quality nonparametrically, there are two population-weighted functions: `popwt_plugin`, which estimates a plug-in estimator of quality, and `popwt_tmle`, which estimates a targeted-minimum-loss-based estimate of quality. The usage for the two functions is similar. The dataset to be used is indicated in the `data` argument, and the columns in the dataset representing observation ID (`id`), the unit identifier (`a`), the outcome (`y`), and covariates to adjust for (`covars`) are indicated with strings.

The output returns the estimates with associated standard errors, as well as shrinkage versions of the estimates under a normal means model. The reliability of the estimate, which measures the relative variance of quality among providers compared to the total variability in the dataset, is also reported. The plug-in results look like this:

```{r}
pop_plugin_est <- popwt_plugin(data = sample_dataset, id = 'id', a = 'unit', y = 'y', covars = c('w1', 'w2', 'w3', 'x1', 'x2', 'x3'))
knitr::kable(pop_plugin_est, digits = 3)
```

The plug-in estimator usage and results are similar:

```{r}
pop_tml_est <- popwt_tmle(data = sample_dataset, id = 'id', a = 'unit', y = 'y', covars = c('w1', 'w2', 'w3', 'x1', 'x2', 'x3'))
knitr::kable(pop_tml_est, digits = 3)
```

### Nuisance function estimation 

These estimators require the estimation of two types of nuisance functions: ones that model the differences between units (propensity scores) and ones that model the relationship between patient characteristics, providers, and the outcome (outcome functions). By default, random forests are used for both types of estimation. One can specify the same learner for all nuisance functions by default through the `lrnr` argument, or separate propensity score (`lrnr_e`) and outcome function (`lrnr_mu`) learners. The `mlr3` package is used for learner specification. Here is an example where multinomial logistic regression is used for propensity scores and standard binary logistic regression is used for the outcome function estimation:

```{r}
pop_plugin_logistic <- popwt_plugin(data = sample_dataset, id = 'id', a = 'unit', y = 'y', covars = c('w1', 'w2', 'w3', 'x1', 'x2', 'x3'), lrnr_e = mlr3::lrn('classif.multinom'), lrnr_mu = mlr3::lrn('classif.log_reg'))
knitr::kable(pop_plugin_logistic, digits = 3)
```

### Equity-weighted quality

The equity-weighted versions of these functions estimate quality in a population of patients where each group defined by the characteristics in `wvars` is given equal weight. In this case, we have three such characteristics:

```{r}
sample_dataset %>%
  count(w1)

sample_dataset %>%
  count(w2)

sample_dataset %>%
  count(w3)
```

The equity-weighted estimators ensure that those with `w3` = 1 are given the same weight as those with `w3` = 0, whereas the population-weighted quality estimates give about 80% weight to those with `w3` = 1. Similarly, those with `w2` = 1 and `w2` = 0 are given the same weight. The usage of these functions is very similar to the population-weighted estimators above, but now one must identify the equity characteristics in the argument `wvars` and the other characteristics to control for in 

```{r}
eq_plugin_est <- eqwt_plugin(data = sample_dataset, id = 'id', a = 'unit', y = 'y', wvars = c('w1', 'w2', 'w3'), zvars = c('x1', 'x2', 'x3'))
knitr::kable(eq_plugin_est %>% select(-wt_ds), digits = 3)
```

The equity-weighted results also include the weights, so users can reason about the equity-weighted population:

```{r}
knitr::kable(eq_plugin_est$wt_ds[[1]], digits = 3)
```
